Vibronic coupling in C$_2$H: implications for astrochemistry

### 2022-03-14 Aug2021-CH2-.tex
update

where the summations run over the dimensionless normal modes ($q_i$) that serve as the coordinate system for the problem.   For convenience, the latter are chosen to be those of the anion, which considerably facilitates calculation of the spectral intensities.  In Eq. 1, the diagonal terms of V (excluding those that carry the Renner-Teller coupling constant $\eta$) represent the quasidiabatic potential energy surfaces of the $\Sigma$, and the two components of the $\Pi$ state, chosen here as those in which the unpaired electron lies within or is perpendicular to an arbitrarily chosen plane, designated as A(a$'$) and A(a$''$), respectively. $\Delta_0^X$ is the separation of the anion and $\Sigma$ states at the origin of the coordinate system (the vertical electron detachment energy), and $\Delta_0^{\Pi}$ is the gap between anion and $\Pi$ states at the same geometry.

For the modes of $\sigma$ symmetry ($q_1$ and $q_3$), the diabatic forces ($F_i$) and force constants ($F_{ij}$) coincide with those of the adiabatic potential energy surfaces.   However, for the bending mode, the diabatic (F$_{22}$) and adiabatic (f$_{22}$) force constants differ, and the parametrization is somewhat more involved.   For each component of the $\Pi$ state, either the 2a or 2b component of the bending vibration will maintain the A$'$ electronic symmetry that is needed to couple with the X state.  Designating these as 2a for the A(a$'$) state and 2b for the A(a$''$) state (as is implicit in Eq. 1), the diabatic force constants for the bending mode in the X and A states can be written as

@@

@@

When the potential in Eq. 1 is diagonalized, the adiabatic states that are used for its parametrization are precisely recovered through second order in displacement. To parametrize this Hamiltonian, the equation-of-motion coupled-cluster method known as EOMIP-CCSDT has been used … (Zach goes from here)…

\begin{equation}
V = \bordermatrix{ & X & A(a') & A(a'') \cr
	X &{\Delta_0^X+F_1^Xq_1+F_3^Xq_3+ 
		\frac{1}{2} \Sigma_{ij}F^X_{ij}q_iq_j}& 0 & 1 \cr
	A(a') & 1 & \Delta^A_0+F^A_1q_1+F^A_3q_3+\frac{1}{2}\sum_{ij}F^A_{ij}q_iq_j+\frac{1}{2}\eta(q_{3a}^2-q_{3b}^2) & 0 \cr
	A(a'') & 0 & 1 & 0 } \qquad
\end{equation}

>>>>

\begin{align}
H &= T_n1+V\\
V &= \bordermatrix{ & X & A(a') & A(a'') \cr
	X &\substack{\Delta_0^X+F_1^Xq_1+F_3^Xq_3+\\ 
		\frac{1}{2} \sum_{ij}F^X_{ij}q_iq_j}& \lambda q_{3a} & \lambda q_{3b} \cr
	A(a') & \lambda q_{3a} & \substack{\Delta^A_0+F^A_1q_1+F^A_3q_3\\+\frac{1}{2}\sum_{ij}F^A_{ij}q_iq_j+\\\frac{1}{2}\eta(q_{3a}^2-q_{3b}^2)} & \eta q_{3a}q_{3b} \cr
	A(a'') & \lambda q_{3b} & \eta q_{3a}q_{3b} & \substack{\Delta_0^A+F_1^Aq_1+F_3^Aq_3\\+\frac{1}{2}\sum_{ij}F_{ij}^Aq_iq_j\\-\frac{1}{2}\eta(q_{3a}^2-q_{3b}^2) } } \qquad
\label{eq:calc1}
\end{align}

where the summations run over the dimensionless normal modes ($q_i$) that serve as the coordinate system for the problem.   For convenience, the latter are chosen to be those of the anion, which considerably facilitates calculation of the spectral intensities.  In Eq.~(\ref{eq:calc1}), the diagonal terms of V (excluding those that carry the Renner-Teller coupling constant $\eta$) represent the quasidiabatic potential energy surfaces of the $\Sigma$, and the two components of the $\Pi$ state, chosen here as those in which the unpaired electron lies within or is perpendicular to an arbitrarily chosen plane, designated as A(a$'$) and A(a$''$), respectively. $\Delta_0^X$ is the separation of the anion and $\Sigma$ states at the origin of the coordinate system (the vertical electron detachment energy), and $\Delta_0^{\Pi}$ is the gap between anion and $\Pi$ states at the same geometry.

For the modes of $\sigma$ symmetry ($q_1$ and $q_3$), the diabatic forces ($F_i$) and force constants ($F_{ij}$) coincide with those of the adiabatic potential energy surfaces.   However, for the bending mode, the diabatic (F$_{22}$) and adiabatic (f$_{22}$) force constants differ, and the parametrization is somewhat more involved.   For each component of the $\Pi$ state, either the 2a or 2b component of the bending vibration will maintain the A$'$ electronic symmetry that is needed to couple with the X state.  Designating these as 2a for the A(a$'$) state and 2b for the A(a$''$) state (as is implicit in Eq.~\ref{eq:calc1}), the diabatic force constants for the bending mode in the X and A states can be written as

\begin{align}
F_{33}^X &= f_{33}^X+\frac{2\lambda^2}{(\Delta_0^A-\Delta_0^X)}\\
F_{33}^A &= f_{3a3a}^{A(a')}-\frac{2\lambda^2}{(\Delta_0^A-\Delta_0^X)}-\eta
\label{eq:calc2}
\end{align}

@@

\begin{equation}
\eta = \frac{1}{2}\left[ f_{3a3a}^{A(a')}-f_{3b3b}^{A(a')}-\frac{2\lambda^2}{(\Delta_0^A-\Delta_0^X)}\right]
\label{eq:calc3}
\end{equation}

@@

When the potential in Eq. 1 is diagonalized, the adiabatic states that are used for its parametrization are precisely recovered through second order in displacement. To parametrize this Hamiltonian, the equation-of-motion coupled-cluster method known as EOMIP-CCSDT has been used to optimise the geometry of the ground state anion, with an Atomic Natural Orbital (ANO) basis (ANO2). Adiabatic linear and quadratic force constants are calculated from the analytical derivatives of the \emph{ab-initio} potential energy surface with respect to the anion reduced normal coordinates, at the anion geometry Q$_0$. The interstate PJT coupling constant $\lambda$ is evaluated directly using the CFOUR package to employ the quasidiabatic ansatz described in refs. This allows for direct ab-initio evaluation of the linear diabatic interstate coupling constants that parametrize the KDC potential. The calculated parameters are presented in the Supplementary Materials, in Table S1.
